---
title: "prac6"
author: "Ruoyu"
date: "2024-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "SMI_P6 Solutions"
output: html_document
---

## Overview

In this practical, we will perform ANOVA and model comparison in R.  
If you would like to install R and R-Studio on your personal device, you can find instructions on MyUni:  
[Installing R and R-Studio](https://myuni.adelaide.edu.au/courses/94623/pages/installing-r-and-r-studio).

In this week’s practical, we will:
- Fit a linear regression model, then perform one-way ANOVA to check for a significant effect.
- Fit identical, parallel, and separate regressions models.
- Select the best model using appropriate criteria.

---

## Load the Data and Libraries

```{r}
library(tidyverse)
library(magrittr)
data("mpg", package = "ggplot2")
mpg
```

---

## Preliminary Analysis

Produce a scatterplot:
```{r}
mpg %>%
  ggplot(aes(x = displ, y = cty)) +
  geom_point()
```

**Question:** What is the relationship between `cty` and `displ`?  
**Solution:** The relationship is negative.

Produce a boxplot:
```{r}
mpg %>%
  ggplot(aes(y = cty, x = drv)) +
  geom_boxplot()
```

**Question:** Which level of `drv` has the highest city fuel efficiency?  
**Solution:** `drv = f` has the highest fuel efficiency.

Produce a scatterplot colored by `drv`:
```{r}
mpg %>%
  ggplot(aes(x = displ, y = cty, colour = drv)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method = lm, se = FALSE)
```

**Question:** Is there a relationship between `drv` and `displ` in predicting `cty`?  
**Solution:** Yes, there appears to be a relationship.

---

## ANOVA

Fit a linear regression model and perform a one-way ANOVA:
```{r}
lm1 <- lm(cty ~ drv, data = mpg)
anova(lm1)
```

**Question:** What is the observed value of the test statistic for the one-way ANOVA?  
**Solution:** $F = 92.676$

**Question:** What are the degrees of freedom?  
**Solution:** $df = 2, 231$

**Question:** Do you accept or reject $H_0$ at $\alpha = 0.05$?  
**Solution:** Reject $H_0$ as the p-value is less than 0.05.

---

## Identical Regressions Model

Fit a linear regression:
```{r}
lm2 <- lm(cty ~ displ, data = mpg)
summary(lm2)
```

**Question:** What is the value of $\hat{\beta_0}$ (intercept)?  
**Solution:** $\hat{\beta_0} = 25.9915$

---

## Parallel Regressions Model

Fit a linear regression:
```{r}
lm3 <- lm(cty ~ displ + drv, data = mpg)
summary(lm3)
```

**Question:** What are the equations for the lines of best fit?  
**Solution:**  
- For `drv = 4`: $cty = 23.6796 - 2.3385 \cdot displ$  
- For `drv = f`: $cty = 25.9547 - 2.3385 \cdot displ$  
- For `drv = r`: $cty = 26.1841 - 2.3385 \cdot displ$

---

## Separate Regressions Model

Fit a regression model with interaction terms:
```{r}
lm4 <- lm(cty ~ displ * drv, data = mpg)
summary(lm4)
```

**Question:** What are the equations for the lines of best fit?  
**Solution:**  
- For `drv = 4`: $cty = 22.5914 - 2.0663 \cdot displ$  
- For `drv = f`: $cty = 28.7198 - 3.4193 \cdot displ$  
- For `drv = r`: $cty = 19.579 - 1.0624 \cdot displ$

---

## Model Selection

Compare models using AIC:
```{r}
AIC(lm2, lm3, lm4)
```

Compare models using BIC:
```{r}
BIC(lm2, lm3, lm4)
```

**Question:** Which model is preferred by AIC?  
**Solution:** The separate regressions model ($AIC = 1051.857$).

**Question:** Which model is preferred by BIC?  
**Solution:** The separate regressions model ($BIC = 1076.045$).

Predict `cty` for a car with 4.7L engine displacement and `drv = 4`:
```{r}
new_car <- data.frame(displ = 4.7, drv = "4")
predict(lm4, newdata = new_car)
```

**Solution:** The predicted value of `cty` is 12.8797.
